---
title: "News Popularity Analysis: Image and Video"
author: "Sushma"
date: "2024-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ROCR)
library(dplyr)
library(ggplot2)
library(corrplot)
```

## Reading data from csv 

```{r}
news_data <- read.csv("/Users/sushmavankayala/Documents/DATA_557/OnlineNewsPopularity/data/OnlineNewsPopularity.csv")
```

## Columns available in the dataset

```{r}
print(colnames(news_data))
```
## Add columns to the dataframe - category, has_image, has_video, num_imgs_std, num_videos_std, is_popular

```{r}
news_data = news_data %>%
  mutate(category = case_when(
    data_channel_is_lifestyle == 1  ~ "Lifestyle",
    data_channel_is_entertainment == 1  ~ "Entertainment",
    data_channel_is_bus == 1  ~ "Business",
    data_channel_is_socmed == 1  ~ "Social Media",
    data_channel_is_tech == 1  ~ "Technology",
    data_channel_is_world == 1  ~ "World",
    TRUE  ~ "Other",
    ),
    has_image = case_when(
      num_imgs > 1 ~ "Yes",
      TRUE ~ "No"
    ),
    has_video = case_when(
      num_videos > 1 ~ "Yes",
      TRUE ~ "No"
    ),
    )

news_data$num_imgs_std = scale(news_data$num_imgs)

news_data$num_videos_std = scale(news_data$num_videos)

news_data$is_popular = news_data$shares > 30000

```

## Checking logistic assumptions - multicolleniarity. Testing between images and videos

```{r}
par(cex = 0.75)

# Assuming house_data is a data frame
data_list <- list(news_data$num_imgs_std, news_data$num_videos_std)

# Convert the list of variables into a matrix and calculate correlations
correlation_values <- cor(do.call(cbind, data_list))

# Add column names to the correlation matrix
colnames(correlation_values) <- c("Number of images(standardized)", "Number of videos(standardized)" )

# Plot the correlation matrix
corrplot(correlation_values, method = "number", bg = "#B4B4B8", number.digits = 2)

```

## Check if presence of images and videos influence popularity of a news article.
### For a logistic regression model fit on the presense of images and popularity, we can do a hypothesis test on the coefficients
```{r}
model_0 = glm(is_popular ~ has_image + has_video, data = news_data)
summary(model_0)
confint(model_0)
```

## Check if number of images and videos influence popularity of a news article.
### For a logistic regression model fit on number of images and popularity, we can do a hypothesis test on the coefficients
```{r}
model_1 = glm(is_popular ~ num_imgs_std + num_videos_std, data = news_data)
summary(model_1)
confint(model_1)
```
## Testing for logistic assumptions

```{r}
ggplot(news_data, aes(x = num_imgs_std, y = predict(model_1, type = "response"))) +
  geom_point() +
  labs(title = "Logit vs. Variable Name", x = "Number of Images(standardized)", y = "Log Odds of Popularity")
```

```{r}
ggplot(news_data, aes(x = num_videos_std, y = predict(model_1, type = "response"))) +
  geom_point() +
  labs(title = "Logit vs. Variable Name", x = "Variable Name", y = "Log Odds of Popularity")
```


## Check if the number of images, number of videos and category influence the popularity of a news article

```{r}
model_2 = glm(is_popular ~ num_imgs_std + num_videos_std + category, data = news_data)
summary(model_2)
confint(model_2)
```

## Check if the number of images, number of videos and category influence the popularity of a news article, considering interaction

```{r}
model_3 = glm(is_popular ~ num_imgs_std * num_videos_std * category, data = news_data)
summary(model_3)
confint(model_3)
```

